Modular Inverse Reinforcement Learning on Human Motion

### 2015-02-11 paper.tex
changing while reading, sec3

Human is able to learn to accomplish complicated tasks much faster than machines
can do. However, most of the tasks can be decomposed into subtasks. A human may
already have the capacity to accomplish the subtasks. He simply transfers the
knowledge of the subtasks and integrate them for the objective task. For example,
when a human learns to drive, he has much prior domain knowledge to help him
with this objective task. He integrates his skills of controlling the velocity,
avoiding other vehicles, navigating, responding to traffic lights, and so on.

@@

A natural question to ask is how human integrates skills for subtasks, and
whether we can use the same method for integrating subtasks in reinforcement
learning problems. In this paper, we analyze human's behavior of
accomplishing a task composite of various subtasks. We collected human motion
data when human accomplish a task, and try to interpret the behavior using
inverse reinforcement learning. With the best of our knowledge, this approach is
noval in the literature.

@@

We will evaluate four different tasks. Task 1, following the path only, and
ignoring other objects. Task 2, following the path, while avoid the obstacles.
Task 3, following the path, while attain targets when possible.  Task 4,
following the path, collecting the targets and avoiding obstacles
simultaneously.
We conducted experiments to ask volunteers to accomplish different tasks and
recorded their trajectories. The human data are collected by the Center for
Perceptual Systems in University of Texas at Austin.

@@

To obtain the weights given the samples, we need to use the Inverse Modular
Reinforcement Learning technique \cite{rothkopf2013modular}. We use a maximum
likelihood method here.
\begin{equation}
\label{eq:irl}
\max_w \prod_t \frac{e^{\eta Q(s^{(t)}, a^{(t)})}}{\sum_b e^{\eta Q(s^{(t)}, b)}}
\end{equation}
where $s^{(t)}$ is the state at time $t$, and $a^{(t)}$ is the action at time
$t$, which are both from samples. $Q(s, a) = \sum_i w_i Q_i(s_i, a)$, as defined
before. $\eta$ is a hyperparameter that determines the consistency of human's
behavior. The larger $\eta$ is, the algorithm more tends to overfit the data.

@@

@@

Weighted sum of Q functions is one way to combine multiple sub-MDPs. We also
propose other ways including, for example, scheduling between different modules,
with only one active at one time. This is also called skilled in the literature
\cite{konidaris2009skill}. However, we adopt the weighted sum approach
because this is more reasonable for human behavior. When a human tries to collect
targets while avoiding obstacles, these two modules are expected to be both
active. A scheduling approach may yield frequent oscillation between these two
modules.

We also assumes independency between modules. Correlation between modules
doesn't impair our analysis in this paper. In Figure~\ref{fig:heatmap}, we can
find that the target module and obstacle module tends to be negatively
correlated.

Weights may be dynamic. However, with such assumption, we need to learn a
mapping from state to weights. In this case, the curse of dimensionality still
exists, and inverse learning would be difficult.

>>>>

Human is able to learn accomplishing complicated tasks much faster than machines
can do. However, most of the tasks can be decomposed into subtasks. A human may
already have the capacity to accomplish the subtasks. He simply transfers the
knowledge of the subtasks and integrate them for the objective task. For example,
when a human learns to drive, he has much prior domain knowledge to help him
with this objective task. He integrates his skills of controlling the velocity,
avoiding other vehicles, navigating, responding to traffic lights, and so on.

@@

A natural question to ask is how human integrates skills for subtasks, and
whether we can use the same method for integrating subtasks in reinforcement
learning problems. In this paper, we analyze human's behavior of
accomplishing a task composite of various subtasks. We collected human motion
data and try to interpret the behavior using a way of
inverse reinforcement learning. With the best of our knowledge, this approach is
novel in the literature.

@@

We will evaluate four different tasks. \bf{Task 1}, following the path only, and
ignoring other objects. \bf{Task 2}, following the path, while avoid the obstacles.
\bf{Task 3}, following the path, while attain targets when possible. \bf{Task 4},
following the path, collecting the targets and avoiding obstacles
simultaneously.
We conducted experiments to ask volunteers to accomplish different tasks and
recorded their trajectories. The human data are collected by the Center for
Perceptual Systems in University of Texas at Austin.

@@

To obtain the weights given the samples, we need to use the Inverse Modular
Reinforcement Learning technique \cite{rothkopf2013modular}. We use a maximum
likelihood method here.
\begin{equation}
\label{eq:irl}
\max_w \prod_t \frac{e^{\eta Q(s^{(t)}, a^{(t)})}}{\sum_b e^{\eta Q(s^{(t)}, b)}}
\end{equation}
where $s^{(t)}$ is the state at time $t$, and $a^{(t)}$ is the action at time
$t$, which are both from samples. $Q(s, a) = \sum_i w_i Q_i(s_i, a)$, as defined
before. $\eta$ is a hyperparameter that determines the consistency of human's
behavior. The larger $\eta$ is, the algorithm is more likely to overfit the data.

@@

From the heatmaps, we find the optimums exist for these tasks. The optimal
weights are also consistent with the experiment context.

@@

We interpreted human behavior using inverse module reinforcement learning in
this paper. We have some positive results, but the performance of our agent is
still inferior than humans.

There are also some observations potential for the future work. First, 
weighted sum of Q functions is one way to combine multiple sub-MDPs. We also
propose other ways including, for example, scheduling between different modules,
with only one active at one time. This is also called skilled in the literature
\cite{konidaris2009skill}. However, we adopt the weighted sum approach
because this is more reasonable for human behavior. When a human tries to collect
targets while avoiding obstacles, these two modules are expected to be both
active. A scheduling approach may yield frequent oscillation between these two
modules.

Second, we also assumes independency between modules. Correlation between
modules doesn't impair our analysis in this paper. In Figure~\ref{fig:heatmap},
we can find that the target module and obstacle module tends to be negatively
correlated.

Lastly, weights may be dynamic and different from state to state. However, with
such assumption, we need to learn a mapping from state to weights. In this case,
the curse of dimensionality still exists, and inverse learning would be
difficult.
